---
title: "analysis_turnover"
author: "Michael Škvrňák"
date: "April 10, 2016"
output: html_document
---

Some charts are [here](http://skvrnami.github.io/). It seems that established political parties rarely nominate a candidate who run on a minor party's list in the previous election, probably because they have enough of their candidates (and it would cause internal clashes...).
New parties (TOP09, VV, ANO etc.) sometimes exploit candidates from minor parties. 
Also, there is a flux of candidates between minor parties. 

```{r, echo=FALSE}
library(ggplot2)
sk1216ord <- read.csv("~/sk1216ord.csv", sep=";")
sk1216ord$party1 <- as.character(sk1216ord$party1)
sk1216ord$party2 <- as.character(sk1216ord$party2)
sk1216ord$select <- sk1216ord$party1==sk1216ord$party2
sk26ord_minor <- sk1216ord[sk1216ord$minor==1, ]
sk26ord_est <- sk1216ord[sk1216ord$minor==0, ]

p <- ggplot(sk26ord_minor, aes(x=order1, y=order2))
p+geom_point(aes(color=select)) + labs(title="Position of candidates in Slovakia, 2012-16, minor parties only", x="candidate's position in 2012", y="candidate's position in 2016", color="Did the candidate change the party?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No"))
print("Correlation:")
cor(sk26ord_minor$order1, sk26ord_minor$order2)

p <- ggplot(sk26ord_est, aes(x=order1, y=order2))
p+geom_point(aes(color=select)) + labs(title="Position of candidates in Slovakia, 2012-16, established parties", x="candidate's position in 2012", y="candidate's position in 2016", color="Did the candidate change the party?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No"))
print("Correlation:")
cor(sk26ord_est$order1, sk26ord_est$order2)
```

```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk1216ord.csv"
turnoverData <- getURL(url)
turnoverData <- read.csv(textConnection(turnoverData))

sk26ord_minor <- turnoverData[turnoverData$minor==1, ]
sk26ord_est <- turnoverData[turnoverData$minor==0, ]
sk26ord_minor$loyal <- as.factor(sk26ord_minor$loyal)

p <- ggplot(sk26ord_minor, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Position of candidates in Slovakia, 2012-16, minor parties only", x="candidate's position in 2012", y="candidate's position in 2016", color="Did the candidate change the party?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No"))
print("Correlation:")
cor(sk26ord_minor$order1, sk26ord_minor$order2)

p <- ggplot(turnoverData, aes(x=order1, y=order2))
p+geom_point(aes(color=improveParty)) + labs(title="Position of candidates in Slovakia, 2012-16", x="candidate's position in 2012", y="candidate's position in 2016", color="Did the candidate move to a better party?") + theme(legend.position="top")

```


```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

table(sk9816ord$loyal, sk9816ord$minorMajor)
table(sk9816ord$loyal, sk9816ord$minorMinor)
table(sk9816ord$loyal, sk9816ord$majorMajor)
table(sk9816ord$loyal, sk9816ord$majorMinor)
```

```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

sk9816ord$loyal <- as.factor(sk9816ord$loyal)
minor <- sk9816ord[sk9816ord$minor==1, ]
major <- sk9816ord[sk9816ord$minor==0, ]

p <- ggplot(minor, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Position of candidates in Slovakia, 2012-16, minor parties", x="candidate's position in 2012", y="candidate's position in 2016", color="Did the candidate change the party?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No"))

p <- ggplot(major, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Position of candidates in Slovakia, 2012-16, major parties", x="candidate's position in 2012", y="candidate's position in 2016", color="Did the candidate change the party?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No"))

```


```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

sk9816ord$loyal <- as.factor(sk9816ord$loyal)
minorMajor <- sk9816ord[sk9816ord$minorMajor==1, ]
# major <- sk9816ord[sk9816ord$minor==0, ]

p <- ggplot(minorMajor, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Candidates improving the status of their party, Slovakia 1998-2016", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate switched between parties?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No")) + geom_abline(intercept = 0, slope=1)

# p <- ggplot(major, aes(x=order1, y=order2))
# p+geom_point(aes(color=loyal)) + labs(title="Position of candidates in Slovakia, 2012-16, major parties", x="candidate's position in 2012", y="candidate's position in 2016", color="Did the candidate change the party?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No"))

table(minorMajor$loyal, minorMajor$improvePosition)

```

```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

sk9816ord$loyal <- as.factor(sk9816ord$loyal)
major <- sk9816ord[sk9816ord$minor==0, ]

p <- ggplot(major, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Candidate from major parties, Slovakia 1998-2016", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate switched between parties?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No")) + geom_abline(intercept = 0, slope=1)

```

#Major => minor
```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

sk9816ord$loyal <- as.factor(sk9816ord$loyal)
majorMinor <- sk9816ord[sk9816ord$majorMinor==1, ]

p <- ggplot(majorMinor, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Candidate from major parties, Slovakia 1998-2016", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate switched between parties?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No")) + geom_abline(intercept = 0, slope=1)

table(majorMinor$loyal, majorMinor$improvePosition)

```

#Minor => minor
```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

sk9816ord$loyal <- as.factor(sk9816ord$loyal)
sk9816ord$origPartyDissolved <- as.factor(sk9816ord$origPartyDissolved)
minorMinor <- sk9816ord[sk9816ord$minorMinor==1, ]

p <- ggplot(minorMinor, aes(x=order1, y=order2))
p+geom_point(aes(color=origPartyDissolved)) + labs(title="Candidates in minor parties", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate's former party stop running in an election?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No")) + geom_abline(intercept = 0, slope=1)

p <- ggplot(minorMinor, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Candidates in minor parties", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate switch between political parties?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No")) + geom_abline(intercept = 0, slope=1)
```

#Minor => minor
```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

sk9816ord$loyal <- as.factor(sk9816ord$loyal)
sk9816ord$origPartyDissolved <- as.factor(sk9816ord$origPartyDissolved)
minorMinor <- sk9816ord[sk9816ord$minorMinor==1, ]

p <- ggplot(minorMinor, aes(x=order1, y=order2))
p+geom_point(aes(color=origPartyDissolved)) + labs(title="Candidates in minor parties", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate's former party stop running in an election?") + theme(legend.position="top") + geom_abline(intercept = 0, slope=1) + scale_color_manual(values=c("black", "red"), labels=c("No", "Yes"))

p <- ggplot(minorMinor, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Candidates in minor parties", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate switch between political parties?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No")) + geom_abline(intercept = 0, slope=1)

table(minorMinor$loyal, minorMinor$improvePosition)

table(minorMinor$origPartyDissolved, minorMinor$improvePosition)

table(minorMinor$improveParty, minorMinor$improvePosition)
```

#Major => major
```{r}
library(RCurl)
library(foreign)
library(ggplot2)
url <- "https://rawgit.com/skvrnami/ma-thesis/master/data_candidates/svk_turnover/sk9816ord.csv"
sk9816ord <- getURL(url)
sk9816ord <- read.csv(textConnection(sk9816ord))

sk9816ord$loyal <- as.factor(sk9816ord$loyal)
sk9816ord$origPartyDissolved <- as.factor(sk9816ord$origPartyDissolved)
majorMajor <- sk9816ord[sk9816ord$majorMajor==1, ]

p <- ggplot(majorMajor, aes(x=order1, y=order2))
p+geom_point(aes(color=loyal)) + labs(title="Candidates in major parties", x="candidate's position in t", y="candidate's position in t+1", color="Did the candidate switch between political parties?") + theme(legend.position="top") + scale_color_manual(values=c("red", "black"), labels=c("Yes", "No")) + geom_abline(intercept = 0, slope=1)

table(majorMajor$loyal, majorMajor$improvePosition)
```


```{r}
model1 <- lm(change ~ order1 + loyal + minor + origPartyDissolved, data=sk9816ord)
summary(model1)
```

```{r}
model2 <- lm(change ~ order1 + loyal*minor + origPartyDissolved, data=sk9816ord)
summary(model2)
```

```{r}
model3 <- lm(change ~ order1 + loyal*minorMajor + loyal*minorMinor + origPartyDissolved, data=sk9816ord)
summary(model3)
```

```{r}
model4 <- lm(change ~ loyal + minorMajor + minorMinor + origPartyDissolved, data=sk9816ord)
summary(model4)
```